---
title: "Homework Assignment 4"
subtitle: "Prioritizing Potential Aquaculture"
author: "Matteo Torres"
date: last-modified
execute:
  output: false
format:
  html:
    theme: darkly
    toc: true
editor_options: 
  chunk_output_type: console
---

### Load Packages
```{r, output = FALSE}
#| code-fold: true

library(here)
library(tmap)
library(stars)
library(terra)
library(tidyverse)
```

### Import Data
```{r}
# West Coast EEZ shapefile
wc_eez <- st_read(here("data", "wc_regions_clean.shp"))

# Bathymetry raster
bathymetry <- rast(here("data", "depth.tif"))

# SST rasters
sst_08 <- rast(here("data", "average_annual_sst_2008.tif"))
sst_09 <- rast(here("data", "average_annual_sst_2009.tif"))
sst_10 <- rast(here("data", "average_annual_sst_2010.tif"))
sst_11 <- rast(here("data", "average_annual_sst_2011.tif"))
sst_12 <- rast(here("data", "average_annual_sst_2012.tif"))

# Raster stack
sst_stack <- c(sst_08, sst_09, sst_10, sst_11, sst_12)
```

```{r, message = TRUE, echo = FALSE}
# Update coordinate reference systems
if ((crs(wc_eez) == crs(bathymetry)) &
    (crs(wc_eez) == crs(sst_stack)) &
    (crs(bathymetry) == crs(sst_stack))) {
  message("All coordinate reference systems match!")
} else {
  bathymetry <- project(bathymetry, crs(wc_eez))
  sst_stack <- project(sst_stack, crs(wc_eez))
  message("All coordinate reference systems are updated and match!")
}
```

### Process Data
```{r}
#| code-fold: true

# Calculate the mean of the stacked raster
mean_sst <- mean(sst_stack, na.rm = TRUE)

# Convert to celsius
mean_sst_celsius <- mean_sst - 273.15

# Crop the depth raster to match SST extent
depth_cropped <- crop(bathymetry, mean_sst_celsius)

# Resample depth raster to match SST resolution
depth_resampled <- resample(depth_cropped, mean_sst_celsius, method = "near")
```

```{r, message = TRUE, echo = FALSE}
if (all.equal(crs(mean_sst_celsius), crs(depth_resampled)) &&
    all.equal(ext(mean_sst_celsius), ext(depth_resampled)) &&
    all.equal(res(mean_sst_celsius), res(depth_resampled))) {
  message("Resolution, extent, and coordinate reference system match!")
} else {
  message("Update the resolution, extent, and coordinate reference system!")
}
```

### Determine the most suitable EEZ
```{r}
#| code-fold: true

# Function to generalize the workflow for any species
identify_suitable_areas <- function(suitable_sst_min, suitable_sst_max, suitable_depth_min, suitable_depth_max, species_name) {
  
  # Reclassify SST: 1 for suitable, 0 for unsuitable
  sst_suitability <- classify(mean_sst_celsius, 
                              rcl = matrix(c(-Inf, suitable_sst_min, 0, 
                                             suitable_sst_min, suitable_sst_max, 1, 
                                             suitable_sst_max, Inf, 0), 
                                           ncol = 3, byrow = TRUE))

  # Reclassify Depth: 1 for suitable, 0 for unsuitable
  depth_suitability <- classify(depth_resampled, 
                                rcl = matrix(c(-Inf, suitable_depth_min, 0, 
                                               suitable_depth_min, suitable_depth_max, 1, 
                                               suitable_depth_max, Inf, 0), 
                                             ncol = 3, byrow = TRUE))

  # Find suitable locations
  suitable_area <- sst_suitability * depth_suitability

  # Mask the suitable area by EEZ boundaries
  suitable_area_eez <- mask(suitable_area, wc_eez)

  # Extract the total suitable area per EEZ
  suitable_area_per_eez <- terra::extract(suitable_area_eez, wc_eez, fun = sum, na.rm = TRUE)

  # Add the result to the EEZ shapefile
  wc_eez$suitable_area <- suitable_area_per_eez[, 2]

  # Visualize the suitable area within each EEZ
  tm_shape(wc_eez) +
    tm_borders() +
    tm_fill("suitable_area", 
            palette = "Blues", 
            title = "Suitable Area (sq km)") +
    tm_layout(main.title = paste("Suitable Area for", species_name, "Aquaculture"),
              main.title.size = 0.6,
              frame = FALSE,
              legend.position = c("right", "center"),
              legend.title.size = 0.8,
              legend.text.size = 0.5) +
    tm_compass(position = c("right", "top")) +
    tm_scale_bar(position = "left")
}
```

#### Oyster Aquaculture
```{r, output = TRUE}
#| code-fold: true

# Oyster conditions
identify_suitable_areas(suitable_sst_min = 11, 
                        suitable_sst_max = 30, 
                        suitable_depth_min = 0, 
                        suitable_depth_max = 70, 
                        species_name = "Oyster")
```

#### White Abalone Aquaculture
```{r, output = TRUE}
#| code-fold: true

# White Abalone conditions
identify_suitable_areas(suitable_sst_min = 10, 
                        suitable_sst_max = 20, 
                        suitable_depth_min = -65, 
                        suitable_depth_max = -6, 
                        species_name = "White Abalone")
```

### Data Citations
|Data|Citation|Link|
|----|--------|----|
|Bathymetry|General Bathymetric Chart of the Oceans. (n.d.). Gridded bathymetry data (general bathymetric chart of the oceans). GEBCO.|https://www.gebco.net/data_and_products/gridded_bathymetry_data/#area|
|Exclusive Economic Zones|Marine regions. (n.d.).|https://www.marineregions.org/eez.php|
|Sea Surface Temperature|Watch, N. C. R. (n.d.). NOAA Coral Reef Watch Daily 5km satellite coral bleaching heat stress SST anomaly product (version 3.1).|https://coralreefwatch.noaa.gov/product/5km/index_5km_ssta.php|
